The problem that will be tackled in this chapter is the ROADEF/EURO Challenge 2020: Maintenance Planning Problem \cite{roadef2020}, which was jointly organized by the French Operational Research and Decision Support Society (ROADEF) and the European Operational Research Society (EURO) in collaboration with RTE (Réseau de Transport d'Électricité).\\

The ROADEF/EURO Challenge is a prestigious biennial competition that bridges the gap between academia and industry by inviting researchers to tackle complex, real-world optimization problems. Since its inception, the challenge has addressed a wide range of industrial issues in collaboration with companies like Renault Group (3D truck loading, 2022), Saint-Gobain (cutting stock optimization, 2018), and Google (machine reassignment in data, 2012). The 2020 edition was motivated by operational necessity and strategic foresight in the energy sector: while preventive maintenance is essential for grid reliability, the act of taking a power line out of service temporarily weakens the network, increasing its vulnerability to unexpected contingencies like extreme weather or other equipment failures. The maintenance optimization problem has become significantly more complex due to the energy transition. The increasing integration of intermittent renewable energies creates new operational dynamics and constraints, making traditional planning methods insufficient. To address this, RTE developed a framework to quantify the financial risk of each potential maintenance task across thousands of future scenarios. The challenge then focused on the second, most complex step: Using this pre-calculated risk data to construct a feasible, robust, and cost-effective annual maintenance schedule.\\

The competition was organized in a systematic multi-phase structure to identify optimal or near-optimal solutions from an international pool of participants. A total of 74 teams from over 20 countries participated across junior (pre-PhD) and senior tracks. The competition proceeded through three main phases, with each phase utilizing different sets of 15 problem instances representing different maintenance planning scenarios. The qualification phase required teams to submit solutions for a public instance set (Set A), with the top 15 teams advancing based on a point-per-instance scoring system. In the semi-final phase, qualified teams tackled a larger instance set (Set B), resulting in 13 finalists (10 senior and 3 junior). The final phase challenged teams with both a known set (Set C) and a hidden set (Set X) of instances. Final rankings were determined through two distinct evaluation runs on the organizers' machines: a 15-minute ``fast score" run and a decisive 90-minute ``quality score" run. The winner was ultimately announced at the EURO 2021 conference. \\


The competition employed a progressive evaluation methodology utilizing increasingly complex instances to assess algorithm effectiveness and scalability. For our purposes, instance difficulty was quantified through objective function scores. The X dataset comprised the most challenging problems, characterized by significant risk penalties during peak demand periods. Therefore, instance \texttt{X12} was selected for detailed analysis despite having only the second-highest objective score after \texttt{X14} which achieved the highest score. The reason is that it demonstrated greater improvement between the 15-minute and 90-minute computational runs. This behavior indicates that extended computation times could generate more diverse scheduling alternatives for our multi-criteria optimization problem.


\section{Original Problem Description}
% The maintenance planning problem aims to schedule the starting times of a set of interventions (i.e., maintenance tasks) over a time horizon (typically one year, discretized into days or weeks). The goal is to minimize risk-related costs represented as a single objective function \ref{eq:objective}, while satisfying resource and exclusion constraints. 
% This is formalized in table \ref{tab:variables} and as an integer programming optimization problem (following the description from \cite{ConsueloRoadef}):\\

The maintenance planning problem aims to schedule the starting times of a set of interventions (i.e., maintenance tasks) over a time horizon (typically one year, discretized into days or weeks). The core of the problem lies in minimizing a composite risk objective while satisfying a set of hard operational constraints.\\

The objective function (Equation \ref{eq:objective}) is a linear convex combination of two distinct risk metrics, balanced by a parameter $\alpha \in [0,1]$. The value of $\alpha$ is not fixed but is provided with each problem instance, allowing the challenge providers to specify the relative importance of each objective and forcing the algorithms to be robust for any value of this parameter. The first term, $\mathrm{obj}_1$, represents the mean cumulative planning risk, driving the schedule toward a low overall average risk. The second term, $\mathrm{obj}_2$, quantifies the ``expected excess", a measure that specifically penalizes schedules with high-risk variability, targeting and discouraging solutions where certain days could face unacceptably high-risk scenarios.\\

This optimization is conditioned by crucial constraints that model real-world limitations. Resource constraints govern the availability of specialized work teams. Each intervention requires a certain number of personnel ($C$), and this consumption ($r_{c,t}$) can fluctuate during the task's duration (e.g., more staff needed for setup and teardown). The model must adhere to both an upper limit ($u_{c,t}$), representing the total teams available, and a lower limit ($l_{c,t}$), ensuring a minimum level of workforce utilization. Furthermore, exclusion constraints address critical safety interdependencies. While the risk of each intervention is calculated individually, certain pairs of tasks are explicitly forbidden from being performed simultaneously. This prevents dangerous synergies not captured in the primary risk model. For example, if performing two specific interventions at the same time would leave a part of the network entirely without supply, they would be forbidden by an exclusion triplet $(i_1, i_2, t)$.\\

This problem is formalized as an integer programming optimization model (following the description from \cite{ConsueloRoadef}), with its primary variables and parameters detailed in Table \ref{tab:variables}:


\begin{table}[!ht]
  \centering
  \begin{tabular}{|c|p{13cm}|}
    \hline
    \textbf{Symbol} & \textbf{Description} \\ \hline
    $\mathcal{T}$ & Total number of time periods in the planning horizon. \\ \hline
    $t$ & Time period index, with $t \in \{1,\ldots,\mathcal{T}\}$. \\ \hline
    $\tau$ & Starting time period for interventions, with $\tau \in \{1,\ldots,\mathcal{T}\}$. \\ \hline
    $\mathcal{I}$ & Set of interventions. \\ \hline
    $x_{i,\tau}$ & Binary variable: 1 if intervention $i\in I$ starts at period $\tau$, 0 otherwise. \\ \hline
    $d_{i,\tau}$ & Duration of intervention $i$ if started at period $\tau$. \\ \hline
    $C$ & Set of resources (teams). \\ \hline
    $l_{c,t},\, u_{c,t}$ & Lower and upper bounds on the availability of resource $c\in C$ at period $t$. \\ \hline
    $r_{c,t}(i,\tau)$ & Consumption of resource $c$ in period $t$ by intervention $i$, if started at $\tau$. \\ \hline
    $\mathcal{E}$ & Set of exclusion triplets $(i_1,i_2,t)$ (interventions $i_1,i_2$ cannot overlap at $t$). \\ \hline
    $S_t$ & Set of scenarios for period $t$. \\ \hline
    $\mathrm{risk}_{s,t}^{(i,\tau)}$ & Risk cost in scenario $s\in S_t$ for intervention $i$ (if started at $\tau$) during period $t$. \\ \hline
    $\alpha$ & Weight in the objective, with $\alpha\in[0,1]$. \\ \hline
  \end{tabular}
  \caption{Main Variables and Parameters}
  \label{tab:variables}
\end{table}

\vspace{1em}


\noindent\textbf{Objective:}
\begin{equation}
\min \; \alpha\,\mathrm{obj}_1 + (1-\alpha)\,\mathrm{obj}_2(\beta) \qquad \text{ with }\alpha \in [0,1]
\label{eq:objective}
\end{equation}
where
\[
\begin{aligned}
&\mathrm{obj}_1 = \frac{1}{\mathcal{T}}\sum_{t=1}^{\mathcal{T}} \mathrm{risk}_t,\quad \mathrm{obj}_2(\beta) = \frac{1}{\mathcal{T}}\sum_{t=1}^{\mathcal{T}} \max\Big\{0,\;Q_t^\beta-\mathrm{risk}_t\Big\},\\[1ex]
&\quad\text{with: }\quad\mathrm{risk}_t = \frac{1}{|S_t|}\sum_{s\in S_t}\sum_{\substack{i\in \mathcal{I}\\ \tau \le t < \tau+d_{i,\tau}}} \mathrm{risk}_{s,t}^{(i,\tau)}\,x_{i,\tau},\quad Q_t^\beta=\beta\text{-quantile of } \Big\{\mathrm{risk}_{s,t} : s\in S_t\Big\}.
\end{aligned}
\]
\noindent Here, $\mathrm{obj}_1$ is the mean cumulative planning risk and $\mathrm{obj}_2(\beta)$ quantifies the extremely high-risk scenarios and periods we want to avoid.

\vspace{1em}
\newpage
\noindent\textbf{Subject to:}

\noindent Each intervention is scheduled exactly once:
\[
\sum_{\tau=1}^{\mathcal{T}-d_{i,\tau}+1} x_{i,\tau} = 1,\quad \forall\, i\in \mathcal{I}
\]

\noindent Intervention finishes within the horizon:
\[
t + d_{i,\tau} \le \mathcal{T}+1,\quad \forall\, i\in \mathcal{I} \text{ and } t \text{ with } x_{i,\tau}=1
\]

\noindent Resource capacity limits:
\[
l_{c,t} \le \sum_{i\in \mathcal{I}}\sum_{\tau \le t < \tau+d_{i,\tau}} r_{c,t}(i,\tau)\,x_{i,\tau} \le u_{c,t},\quad \forall\, c\in C,\; \forall\, t
\]

\noindent Exclusion constraints:
\[
\sum_{\tau:\, t\in [\tau,\,\tau+d_{i_1,\tau}-1]} x_{i_1,\tau} + \sum_{\tau:\, t\in [\tau,\,\tau+d_{i_2,\tau}-1]} x_{i_2,\tau} \le 1,\quad \forall\,(i_1,i_2,t)\in \mathcal{E}
\]

\noindent Binary variables:
\[
x_{i,\tau} \in \{0,1\},\quad \forall\, i\in \mathcal{I},\; \tau=1,\ldots,\mathcal{T}-d_{i,\tau}+1
\]


