In the literature\cite{HistoryFL2017}, fuzzy logic is understood in 2 different ways: broad ($\FLb$) or narrow ($\FLn$, also called mathematical fuzzy logic). The former is regarding Zadeh's agenda (more practical) and the latter is more related to the formal logics that can be derived from fuzzy sets. However, this section only aims to very briefly introduce without proof some of the main results and concepts related to $\FLn$: \signal{fuzzy implications, different fuzzy logics derived from t-norms and approximate reasoning. HAJEK DICE QUE NO ESTÁ ACABADO EL FLn, VER POR QUÉ. PAG 14PDF}\\

Some books (such as \cite{HistoryFL2017}) first present the departure from classical bivalence of truth values to the generalization of degrees of truth. Then arises the equivalence between truth values of fuzzy predicates like "This car is expensive", and the membership of that car to the fuzzy set that represents the concept of expensive. However in this work, the inverse order has been followed as it is more familiar to someone with a background in mathematics. The next two paragraphs aim to illustrate the relationship between set operations and their logical counterparts, which will be used for defining fuzzy logical connectives.\\


In classical propositional logic\footnote{See appendix \ref{app:form_log} for a brief introduction to the concepts of formal logic}, there is a very straightforward relation between set-membership and truth values. It is the same to state \say{$x$ belongs to $A$} or \say{(it is true that) $x$ is $A$}. This direct correspondence extends to the fundamental operations of sets and logical connectives: stating that \say{$x$ belongs to the intersection of $A$ and $B$} (denoted $x \in A \cap B$) is equivalent to asserting that \say{$x$ is $A$ AND $x$ is $B$} (the logical conjunction $P_A(x) \land P_B(x)$ is true). The case for union $\cup$ and disjunction $\lor$ is analogous. Finally, the concept of an element $x$ belonging to the negation of set $A$ ($x \in \overline{A}$), meaning $x$ does not belong to $A$, directly mirrors logical negation ($\neg$). When working with fuzzy sets, we consider partial truth (degrees of truth as degrees of membership).\\

The implication connective is rooted in the residuation (or adjointness) property, which views implication $I$ ($A \Rightarrow B$, or "if $A$, then $B$") as a form of logical division: $A \land I \models B$ in logic (or $A \cap I \subseteq B$ in set theory). The goal is to find the weakest proposition (or largest set)\footnote{Intersection with a larger set $I$ is less restrictive respect to inclusion of sets. If intersection with the largest set is contained, intersection with any smaller set will also be contained.} $I$ such that combining $A$ with $I$ through conjunction (intersection) yields something at least as strong as (contained in) $B$. This largest $I$ satisfying the condition defines the implication $A \Rightarrow B$, internalizing the notion of logical consequence. Logical equivalence ($A \iff B$, or "$A$ if and only if $B$") occurs when both $A \Rightarrow B$ and $B \Rightarrow A$ hold. In set theory, this corresponds to mutual subset relations ($A \subseteq B$ and $B \subseteq A$), which can only be true when sets $A$ and $B$ are identical. Thus, logical equivalence between propositions directly corresponds to set equality.\\


When extending the notion of truth from classical to fuzzy logic, two fundamental questions arise: why the unit interval $[0,1]$ as truth value set and what is the interpretation of specific truth degrees \cite[Sec.~1.1]{Hajek1998}. The unit interval provides a continuous scale accommodating any granularity while maintaining a linear order ($\leq$) with 1 representing absolute truth and 0 representing absolute falsity. This generalizes finite truth value sets and aligns with real world scenarios where responses can be normalized to $[0,1]$. Furthermore, when restricting to the subset $\{0,1\}$, fuzzy logic must reduce to classical two valued logic as a special case, ensuring backward compatibility with established logical frameworks.\\

The meaning of truth degrees differs between atomic and compound propositions. For atomic propositions, truth degrees represent direct evaluations (e.g., a response of "more or less" to "Do you like Haydn?" maps to 0.7). Importantly, truth degrees cannot be meaningfully compared across different propositions (or it would lead to illogical comparisons such as "You like Haydn more than you are old"). For compound propositions, most fuzzy logics are truth functional: the truth degree of a complex formula is determined by a function of its components' truth degrees, similar to how classical truth tables.\\



\paragraph{Conjunction, Disjunction, and Negation}\signal{Esto lo puedo acortar yo creo}In classical logic, conjunction corresponds to set intersection. When generalizing to fuzzy sets, the truth function for conjunction should capture the intuition that \say{a large truth degree of $\phi \text{ \& } \psi$ should indicate that both the truth degree of $\phi$ and the truth degree of $\psi$ is large}\cite[p.27]{Hajek1998}. This leads to the adoption of \textbf{t-norms} (triangular norms) as the truth functions for fuzzy conjunction. The three basic continuous t-norms (\luka, Gödel/minimum, and Product) are particularly important as they give rise to three fundamental fuzzy logics: Lukasiewicz logic ($\L$), Gödel logic ($G$) and Product logic ($\Pi$), respectively. \signal{MENCIONAR QUÉ AXIOMAS SON ESOS QUE LOS HACEN ESPECIALES Y SI EL HECHO DE QUE ESAS T-NORMAS GENEREN TODAS LAS NORMAS CONTINUAS, TIENE ALGUNA IMPLICACIÓN. Each of these logics is characterized by specific axioms that reflect the algebraic properties of its underlying t-norm, making them cornerstones in the study of fuzzy logical systems.}

Fuzzy \textbf{disjunction (OR)} and fuzzy \textbf{negation (NOT)} are generalized from their classical set-theoretic counterparts (union and complement, respectively) and are typically defined using t-conorms and fuzzy complement functions, as detailed in Section~\ref{sec:set_operations}.

\subsection{Fuzzy Implications}

\signal{Falta entender bien el concepto de implicación material. Lo que se entiende por implicación material (propiedad del residuo) y lo que no y qué repercusiones tiene eso para la lógica.}

The implication connective ($P \rightarrow Q$, "IF $P$ THEN $Q$") is arguably the most critical for logical reasoning and building rule-based systems, such as those used to model expert opinion in MCDM, a topic that will be explored further in the next chapter. Its generalization to fuzzy logic presents a rich variety of approaches. A fuzzy implication $I(x,y)$ (where $x$ is the truth value of the antecedent and $y$ that of the consequent) aims to quantify "to what extent $q$ is at least as true as $p$" (\cite[p.57]{FULLER2}). Different formalizations capture this intuition in different ways. The choice among them depends heavily on the specific application and the desired properties.

\paragraph{R-Implications (Residuated Implications)}
This class of implications arises directly from the residuation principle you introduced earlier and is fundamental in algebraic fuzzy logic. Given a t-norm $T$ (representing conjunction), its corresponding R-implication $I_R$ is defined as the residuum of $T$:
\[I_R(x,y) = \sup \{z \in [0,1] \mid T(x,z) \le y\}\]
This definition ensures that $I_R(x,y)$ is the weakest (largest) truth value $z$ such that when $x$ is conjoined with $z$ (using $T$), the result is no stronger (no greater) than $y$ (\cite[Def. 2.1.5, p.29]{Hajek1998}). This property is crucial for the soundness of modus ponens in fuzzy logic. \signal{Esto se puede extendar a left semicont.}
\begin{example}[R-Implications from Continuous T-norms \cite{Hajek1998}]
The main continuous t-norms yield the following R-implications:
\begin{itemize}
    \item The \textbf{Łukasiewicz Implication} (from $T_L(x,y) = \max(0, x+y-1)$):
    \[I_L(x,y) = \min(1, 1-x+y)\]
    \item The \textbf{Gödel Implication} (from $T_G(x,y) = \min(x,y)$):
    \[I_G(x,y) = \begin{cases} 1 & \text{if } x \le y \\ y & \text{if } x > y \end{cases}\]
    \item The \textbf{Goguen (or Product) Implication} (from $T_P(x,y) = xy$):
    \[I_P(x,y) = \begin{cases} 1 & \text{if } x \le y \\ y/x & \text{if } x > y \end{cases}\]
\end{itemize}
\end{example}
R-implications are vital for constructing axiomatic fuzzy logical systems because they possess strong algebraic properties that ensure logical coherence. \signal{Explicar esto más.}

\paragraph{S-Implications (Standard Implications)}
This family of implications is derived from the classical logical equivalence $P \rightarrow Q \equiv \neg P \lor Q$. A fuzzy S-implication is defined using a fuzzy negation and a t-conorm $S$ \cite[p.59]{FULLER2}:
\[I_S(x,y) = S(\overline{x}, y)\]
\begin{example}[S-Implications]
Common examples include:
\begin{itemize}
    \item The \textbf{Kleene-Dienes Implication}, using standard negation $\overline{x}=1-x$ and maximum t-conorm $S(a,b)=\max(a,b)$:
    \[I_{KD}(x,y) = \max(1-x, y)\]
    \item If the Łukasiewicz t-conorm $S_L(a,b) = \min(1, a+b)$ is used with standard negation, the resulting S-implication is $I_L(x,y) = \min(1, 1-x+y)$, which is identical to the Łukasiewicz R-implication.
\end{itemize}
\end{example}

\paragraph{Implications from T-norms (for Fuzzy Rules)}
In many practical applications, especially in fuzzy control and Multiple-Input Single-Output (MISO) fuzzy systems used in MCDM, implications are defined directly using a t-norm:
\[I(x,y) = T(x,y)\]
\begin{example}[T-norm based Implications \cite{FULLER2}]
Popular examples include:
\begin{itemize}
    \item The \textbf{Mamdani Implication}: $I_M(x,y) = \min(x,y)$ (using the Gödel t-norm).
    \item The \textbf{Larsen Implication}: $I_{La}(x,y) = xy$ (using the Product t-norm).
\end{itemize}
\end{example}
It is crucial to recognize that, as \cite[p.60]{FULLER2} notes, ``these implications do not verify the properties of material implication'' in all respects. For instance, with the Mamdani implication, $I_M(0,0) = \min(0,0) = 0$, whereas classical and most other fuzzy implications yield $I(0,0)=1$. However, in the context of fuzzy rule-based systems, if an antecedent is false (truth value 0), the rule often contributes nothing to the output, making this behavior desirable.

\paragraph{Standard Strict Implication}
The most direct generalization of the classical idea that $P \rightarrow Q$ is true if and only if the truth value of $P$ is less than or equal to the truth value of $Q$ is the \textbf{Standard Strict Implication} (\cite[p.58]{FULLER2}):
\[I_{SS}(x,y) = \begin{cases} 1 & \text{if } x \le y \\ 0 & \text{if } x > y \end{cases}\]
While intuitive, this implication is highly sensitive to small changes in input values (e.g., $I_{SS}(0.8, 0.8)=1$ but $I_{SS}(0.8, 0.7999)=0$), making it less suitable for applications dealing with imprecise numerical data \cite[p.58]{FULLER2}.

 

































\subsection{Fuzzy implications}
S-implications
R-implications
t-norm implications
\signal{$(\fuzzy{X},\cup , \cap , \lnot)$ is a complete, completely distributive, lattice with an involution. This extends the boolean algebra.}

\signal{The notion of equality is replaced by a graded relation (often measured via the biresiduum $\leftrightarrow$)}

\signal{Fuzzy description logics explained %https://www.umbertostraccia.it/cs/download/papers/KES09/KES09.pdf

The Lukasiewicz t-conorm is closely related to the basic binary operation of multi-valued
algebras. Additionally, t-norms and t-conorms form examples of aggregation operators. They
play a significant role in the axiomatic definition of the concept of triangular norm-based measure
and, in particular, of the concept of probability of fuzzy events; the Frank family of t-norms and
t-conorms plays a particular role [6]. 

%https://cake.fiu.edu/Publications/Ngan+al-18-LC.Logic_Connectives_of_Complex_Fuzzy_Sets_ROMJIST_downloaded.pdf 

It should be mentioned that t-norms overlap with copulas [3, 24]: commutative associative
copulas are t-norms; t-norms which satisfy the 1-Lipschitz condition are copulas. Some families
of t-norms are known as families of copulas under different names}

\signal{
    Lukasiewicz logic algebraic structure and formal implications thanks to the residual property.\\

    Gödel Logic uses 
    $a \otimes b = \min(a, b)$, which tends to produce conservative scores dominated by the weakest criterion. While useful for risk-averse scenarios, it fails to differentiate alternatives when all criteria are partially satisfied.

    Product Logic employs 
    $a \otimes b = a \cdot b$, amplifying the impact of low-scoring criteria. This can lead to premature elimination of alternatives with one poor attribute.

    Łukasiewicz Logic's additive t-norm balances compensation between criteria, allowing alternatives to offset weaknesses in one dimension with strengths in others—a critical feature for complex trade-off analysis.
}








\subsection{Approximate Reasoning and Pavelka-Style Completeness}

Classical deductive systems typically deal with absolute truth: premises are assumed to be fully true (truth value 1), and sound inference rules guarantee that conclusions are also fully true. However, in many real-world scenarios and in the spirit of fuzzy logic, we often reason with information that is only partially true or true to a certain degree. This leads to the field of **approximate reasoning**, where we are interested in the degree to which a conclusion follows from premises that themselves have degrees of truth.

A significant contribution to formalizing reasoning with degrees of truth was made by Jan Pavelka in the 1970s, particularly for logics related to Łukasiewicz logic. His work was later refined and integrated into the broader t-norm based framework by Hájek. The core idea is to extend a fuzzy logical system (like Łukasiewicz logic, denoted L) with rational truth constants $\bar{r}$ for every rational $r \in [0,1] \cap \mathbb{Q}$. A formula $(\bar{r} \rightarrow \phi)$ can then be interpreted as stating "$\phi$ is true to a degree at least $r$".

This framework allows for the definition of two key concepts for a given theory $T$ (a set of axioms, each potentially associated with a truth degree or assumed to be fully true):

\begin{itemize}
    \item \textbf{Truth Degree of $\phi$ over $T$}: $||\phi||_T = \inf\{e(\phi) \mid e \text{ is a model of } T\}$. This is the lowest degree to which $\phi$ is true in all models that satisfy the theory $T$.
    \item \textbf{Provability Degree of $\phi$ over $T$}: $|\phi|_T = \sup\{r \mid T \vdash (\bar{r} \rightarrow \phi)\}$. This is the highest degree $r$ such that it is provable from $T$ that $\phi$ is true to degree at least $r$.
\end{itemize}

Pavelka's completeness theorem for Rational Pavelka Logic (RPL), which is Łukasiewicz logic extended with rational truth-value constants and appropriate "bookkeeping" axioms for these constants, establishes a fundamental connection between these semantic and syntactic degrees.

\begin{theorem}[Pavelka-style Completeness for RPL]
For any theory $T$ in Rational Pavelka Logic and any formula $\phi$:
\[ |\phi|_T = ||\phi||_T \]
\end{theorem}

This remarkable result states that the highest degree to which a formula is provable from a theory $T$ is precisely the lowest degree to which it is true in all models of $T$. It provides a formal calculus for "approximate reasoning" where conclusions can be derived with associated degrees of truth based on the degrees of truth of the premises. For instance, if we have premises $A_1, \dots, A_k$ with associated truth degrees $r_1, \dots, r_k$, and we can syntactically derive a conclusion $C$ with an associated provability degree $s$ that depends on $r_1, \dots, r_k$, Pavelka's theorem ensures that this syntactically derived degree $s$ is the "best possible" semantic guarantee for the truth of $C$. While the original RPL is specific to Łukasiewicz logic due to its reliance on the properties of the Łukasiewicz t-norm and its residuum, the general concept of graded provability and its relation to graded truth is a cornerstone of advanced fuzzy logic.